Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-16 manuscript/manuscript.tex
update ms.

Why does this matter for research in psychology? Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the timecourse of psychological states. As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
For example, the distributions in Example 4 show that the effect starts around 200 ms and is gone by 600 ms. In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms. And in the Example 6, the effect reverses around between 500 and 600 ms. What kind of theory or set of theories could account for such effects? Are there new auxiliary assumptions that theories need to adopt? And are there new experiments that need to be run to test the novel predictions that follow from these analyses? As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to theory building toolkit.

@@

In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (REFs) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.

@@

For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs). Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs). Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before. Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology. As such, we only supply regression equations in supplementary materials and then refer to them in the text whenever relevant.

@@

The definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. Since our focus here is on hazard models that use discrete time units, we describe that approach. After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition. For example, the first response could occur in 550 ms and it would be in timebin 6 (any RTs from 501 ms to 600).

Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X). The equations that define both of these functions are reported in supplementary materials (Supp XX). The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in a subsequent timebin.

@@

Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013). This means that wherever possible we include varying intercepts and slopes per pid.
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

{[}{[}provide a short summary of the aims and scope of each tutorial, as well as the links between them{]}{]}. Additionally, to further simplify the process for other users, the tutorials rely on a set of our own user-defined functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.

@@

\subsection{4.1 Tutorial 1: Calculating descriptive statistics using a life table}\label{tutorial-1-calculating-descriptive-statistics-using-a-life-table}

@@

Our data wrangling procedures serve two related purposes. First, we want to summarise and visualise descriptive statistics that relate to our main research questions. Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches. The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2). The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment. In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred. It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest. Likewise, the event could occur in bin 1 there would only be 1 row of data for that trial.

@@

\subsection{4.2 Tutorial 2: Fitting Bayesian hazard models}\label{tutorial-2-fitting-bayesian-hazard-models}

@@

\subsubsection{4.2.1 Hazard model considerations}\label{hazard-model-considerations}

@@

\subsubsection{4.2.2 Prior distributions}\label{prior-distributions}

@@

\subsubsection{4.2.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}

@@

\subsubsection{4.2.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}

@@

\subsubsection{4.2.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}

@@

\subsubsection{4.2.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}

@@

\subsubsection{4.2.7 Compare the models.}\label{compare-the-models.}

@@

\subsubsection{4.2.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}

@@

\subsection{4.3 Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}

@@

\subsection{4.4 Tutorial 4: Generalising to a more complex design}\label{tutorial-4-generalising-to-a-more-complex-design}

So far in this paper, we have use a simple experimental design, which involved one condition with two levels. But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels. The purpose of Tutorial 4, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design. We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.

To this end, Tutorial 4 illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type. As we use the same functional programming approach as in Tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8. Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.

{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_4_descriptive_2IV/Plot_2IV_for_subject6_PanisSchmidt}

@@

Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution. Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function. Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom). For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect? After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions. Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).

@@

@@

\section{Supplementary Material}\label{supplementary-material}

here is the supp

>>>>

Why does this matter for research in psychology? Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the timecourse of psychological states. As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
For example, the distributions in Example 4 show that the effect starts around 200 ms and is gone by 600 ms. In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms. And in the Example 6, the effect reverses around between 500 and 600 ms. What kind of theory or set of theories could account for such effects? Are there new auxiliary assumptions that theories need to adopt? And are there new experiments that need to be run to test the novel predictions that follow from these analyses? As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to our theory building toolkit.

@@

In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.

@@

For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (Singer \& Willett, 2003). Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs). Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before. Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology. As such, we only supply regression equations in supplementary materials and then refer to them in the text whenever relevant.

@@

The definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. Since our focus here is on hazard models that use discrete time units, we describe that approach. After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition. For example, the first response could occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600). Continuous RT data is treated here as interval-censored data.

Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X). The equations that define both of these functions are reported in supplementary materials (part A). The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occured in the past, i.e., in one of the prior bins. This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.

@@

Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013). This means that wherever possible we include varying intercepts and slopes per participant
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

Tutorials 1a and 1b show how to calculate and plot the descriptive statistics when there are one and two independent variables, respectively. Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
Tutorials 3a and 3b show how to implement multilevel models for hazard and conditional accuracy in the frequentist framework, respectively.
Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.

@@

\subsection{4.1 Tutorial 1a: Calculating descriptive statistics using a life table}\label{tutorial-1a-calculating-descriptive-statistics-using-a-life-table}

@@

Our data wrangling procedures serve two related purposes. First, we want to summarise and visualise descriptive statistics that relate to our main research questions. Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches. The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2). The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment. In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred. It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest. Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.

@@

\subsection{4.2 Tutorial 1b: Generalising to a more complex design}\label{tutorial-1b-generalising-to-a-more-complex-design}

So far in this paper, we have use a simple experimental design, which involved one condition with two levels. But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels. The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design. We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.

To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type. As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8. Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.

{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt}

\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models}\label{tutorial-2a-fitting-bayesian-hazard-models}

@@

\subsubsection{4.3.1 Hazard model considerations}\label{hazard-model-considerations}

@@

\subsubsection{4.3.2 Prior distributions}\label{prior-distributions}

@@

\subsubsection{4.3.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}

@@

\subsubsection{4.3.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}

@@

\subsubsection{4.3.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}

@@

\subsubsection{4.3.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}

@@

\subsubsection{4.3.7 Compare the models.}\label{compare-the-models.}

@@

\subsubsection{4.3.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}

@@

\subsection{4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}

\subsection{4.5 Tutorial 3a: Fitting Frequentist hazard models}\label{tutorial-3a-fitting-frequentist-hazard-models}

@@

\subsection{4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models}\label{tutorial-3b-fitting-frequentist-conditional-accuracy-models}

@@

Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.

Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution. Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function. Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom). For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect? After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions. Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).

We did not discuss continuous-time hazard analysis. As indicated by --allison -- learning discrete-time methods first will help in learning continuous-time methods. Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail. However, they require much more data to estimate the continuous-time hazard (rate) function well. Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.

@@

\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.

@@

\section{Supplementary material}\label{supplementary-material}

\subsection{A. Definitions of discrete-time hazard, surivor, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-and-conditional-accuracy-functions}

The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991). After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
Discrete-time EHA focuses on the discrete-time hazard function

\noindent h(t) = P(RT = t\textbar{} RT \(\geq\) t) \hfill  (1)

\noindent and the discrete-time survivor function

\noindent S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} \hfill  (2)

\noindent and not on the probability mass function

\noindent P(t) = P(RT = t) = h(t).S(t-1) \hfill  (3)

\noindent nor the cumulative distribution function

\noindent F(t) = 1-S(t) = P(RT \(\leq\) t). \hfill  (4)

The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins. Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.

For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function

\noindent ca(t) = P(correct \textbar{} RT = t) \hfill  (5)

which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977). This latter function is also known as the micro-level speed-accuracy tradeoff function.

When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t). The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.

\subsection{B. Custom functions for descriptive discrete-time hazard analysis}\label{b.-custom-functions-for-descriptive-discrete-time-hazard-analysis}

We defined 13 custom functions that we list here.

\begin{itemize}
\tightlist
\item
  censor(df,timeout,bin\_width)
\item
  ptb(df)
\item
  setup\_lt(ptb)
\item
  setup\_lt\_2IV(ptb)
\item
  calc\_ca(df)
\item
  calc\_ca\_2IV(df)
\item
  join\_lt\_ca(df1,df2)
\item
  join\_lt\_ca\_2IV(df1,df2)
\item
  extract\_median(df)
\item
  extract\_median\_2IV(df)
\item
  plot\_eha(df,subj,haz\_yaxis)
\item
  plot\_eha\_2IV(df,subj,haz\_yaxis)
\item
  plot\_eha\_agg(df,subj,haz\_yaxis)
\end{itemize}

\subsection{C. Link functions}\label{c.-link-functions}

{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}

\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
\end{figure}

\subsection{D. Regression equations}\label{d.-regression-equations}

An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.

The main predictor variable TIME is the time bin index t that is centered on value 9 in this example. The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero). The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable. Predictors can be discrete, continuous, and time-varying or time-invariant. For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units. However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth. To interpret the effects of a predictor,its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link). When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.

An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.

\subsection{E. Prior distributions}\label{e.-prior-distributions}

To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution. Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale. According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).

{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}

\caption{Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
\end{figure}

To gain a sense of what prior \emph{cloglog} values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function). Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
